A Tale of Two Transects

### 2019-11-20 writing/transect_paper.Rmd
writing methods

```{r setup, include=FALSE}
library("knitr")
library("tikzDevice")
knitr::opts_chunk$set(dev="tikz",
                      # dev.args = list(pointsize=32),
                      echo=FALSE,
                      results='asis',
                      fig.show='asis',
                      external=TRUE)
options(tikzDefaultEngine = "luatex",
        tikzLualatexPackages = c(options("tikzXelatexPackages")$tikzXelatexPackages,
                                "\\usepackage[version=4]{mhchem}\n",
                                "\\usepackage{siunitx}\n",
                                "\\DeclareSIUnit\\year{a}\n"))

@@

## DNA extraction, PCR, and sequencing

After field lysis, DNA was extracted using the Zymo Research Soil/Fecal Prep kit **check name**.
DNA for short amplicons was amplified using the primer pairs gITS7[@ihrmark2012]--ITS4[@white1990amplification].
  - short amplicon : gITS7 -- ITS4 (multiplexing index on gITS7)
  - short PCR protocol:
  - long amplicon : ITS1 -- LR5 (multiplexing index on both ends)
  - long PCR protocol:
  - short amplicon : same library sequenced with PacBio RS II and Ion Proton S5
  - long amplicon : PacBio RS II

>>>>

```{r setup, include=FALSE}
library("knitr")
library("tikzDevice")
tryCatch(
  tinytex::tinytex_root(),
  error = function(e) tinytex::install_tinytex()
  )
knitr::opts_chunk$set(dev="tikz",
                      # dev.args = list(pointsize=32),
                      echo=FALSE,
                      results='asis',
                      fig.show='asis',
                      external=TRUE)
options(tikzDefaultEngine = "luatex",
        tikzLualatexPackages = c(options("tikzXelatexPackages")$tikzXelatexPackages,
                                "\\usepackage[version=4]{mhchem}\n",
                                "\\usepackage{siunitx}\n",
                                "\\DeclareSIUnit\\year{a}\n"))

@@

## DNA extraction, amplification, and sequencing

After field lysis, DNA was extracted using the Zymo Research Soil/Fecal Prep kit **check name**.
DNA was quantified using PicoGreen *according to the manufacturer's protocol*?

Two different regions of rDNA were amplified (figure).
The short amplicon (approximately 300 bp) included the full ITS2 region as well as parts of the flanking 5.8S and LSU rDNA, using gITS7 [@ihrmark2012] as the forward primer and a mix of ITS4 [@white1990amplification] and ITS4A [@larena1999] as the reverse primer.

DNA for short amplicons was amplified using gITS7 [@ihrmark2012] as the forward primer and a equimolar mix of ITS4 [@white1990amplification] and ITS4A [@larena1999] as the reverse primer (hereafter ITS4m).
gITS7 primers were indexed for multiplexing (**supp info**).
Amplification was performed by polymerase chain reaction (PCR) in 60µl reactions containing 200 µM dNTP mix, 250 µM indexed gITS7 primer, 150 µM ITS4m, 2 mM McCl2, 0.3 U *taq* polymerase (Dream *Taq*, Thermo Fisher) and 10--20 ng purified DNA in Dream *Taq* buffer.
The reaction conditions were 10 min at 95°, followed by 35 cycles of 60 s at 95°, 45 s at 56°, and 50 s at 72°, and finally 3 min at 72°.
Each reaction was split into three aliquots and amplified separately to reduce the effect of PCR stochasticity (**ref**), and then pooled after amplification.

```{tikz rDNA_amplicons, fig.cap = "rDNA amplicons", code = readLines("rDNA_amplicons.pgf"), fig.ext = "pdf"}

```

DNA for long amplicons was extracted using ITS1 [@white1990amplification] for the forward primer and LR5 [@vilgalys1990] as the reverse primer.
Both primers were indexed for multiplexing (**supp info**).
PCR was performed in 60 µL reactions as for the short amplicons, but with 500 µM of each of the two primers.
Reaction conditions were 10 min at 95°, 30 cycles of 45 s at 95°, 45 s at 59°, and 90 s at 72°, and finally 10 min at 72°.
Each reaction was split into three aliquots as for short amplicons.

  - short amplicon : same library sequenced with PacBio RS II and Ion Proton S5
  - long amplicon : PacBio RS II
